---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

# **`conftree`**: Subgroups With Regression Trees Using Conformal Prediction

This package uses conformal prediction for subgroup detection with regression trees. It is based on the `r2p` algorithm, as introduced by [Lee et. al (NeurIPS, 2020)][1]. In `conftree`, we improve this framework by extending it to [CV+/Jackknife+][2]. This works with any regression model that can be fitted with `tidymodels`.

## Installation

You can install the current development version from GitHub with:

```{r gh-installation, eval = FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("holgstr/conftree")
```

## Quickstart

Let's find subgroups in the Washington bike share data. We use `tidymodels` to set a random forest as `learner`, a 5% miscoverage rate as `alpha`, and 10-fold CV+ to quantify the uncertainty in the resulting subgroups:

```{r, message = FALSE, eval = FALSE}
library(conftree)
library(tidymodels)
data(bikes)

# Specify the learner used for model training:
set.seed(1234)
randforest <- rand_forest(trees = 200, min_n = 5) %>%
  set_mode("regression") %>%
  set_engine("ranger")

# Find optimal subgroups using conformal prediction:
groups <- r2p(data = bikes,
              target = "count",
              learner = randforest,
              cv_folds = 10,
              alpha = 0.05,
              gamma = 0.2,
              lambda = 0.5)

# Display tree structure:
groups$tree

```

```{r, message = FALSE, echo = FALSE}
library(tidymodels)
devtools::load_all()
data(bikes)

# Specify the learner used for model training:
set.seed(1234)
randforest <- rand_forest(trees = 200, min_n = 5) %>%
  set_mode("regression") %>%
  set_engine("ranger")

# Find optimal subgroups using conformal prediction:
groups <- r2p(data = bikes,
              target = "count",
              learner = randforest,
              cv_folds = 10,
              alpha = 0.05,
              gamma = 0.2,
              lambda = 0.5)

# Display tree structure:
groups$tree

```

[1]: https://proceedings.neurips.cc/paper/2020/hash/1819020b02e926785cf3be594d957696-Abstract.html

[2]: https://arxiv.org/abs/1905.02928
