---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# **`conftree`**: Subgroups with Trees using Conformal Prediction

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R-CMD-check](https://github.com/holgstr/conftree/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/holgstr/conftree/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This experimental `R` package detects subgroups in machine learning models with numeric outcomes (regression). It leverages [CV+/Jackknife+][1] to simultaneously optimize inter-group heterogeneity and intra-group homogeneity w.r.t. the predicted target variable. In this way, it is an extension of the `r2p` algorithm ([Lee et al., NeurIPS 2020][2]). `conftree` works with any of the [100+][3] learning algorithms for regression tasks available in `tidymodels`.

## Installation

You can install the current development version from GitHub with:

```{r gh-installation, eval = FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("holgstr/conftree")
```

## Quickstart

Let's find subgroups in the Washington bike share data. We use `tidymodels` to set a random forest as `learner`, a 5% miscoverage rate as `alpha`, and 10 `cv_folds` for the CV+ to quantify the uncertainty in the resulting subgroups:

```{r, message = FALSE}
library(conftree)
library(tidymodels)
data(bikes)

set.seed(1234)

# Specify the learner to be used for model training:
randforest <- rand_forest(trees = 200) %>%
  set_mode("regression") %>%
  set_engine("ranger")

# Find optimal subgroups:
groups <- r2p(
  data = bikes,
  target = "count",
  learner = randforest,
  cv_folds = 10,
  alpha = 0.05,
  gamma = 0.2,
  lambda = 0.5,
  max_groups = 4
)

# Display tree structure:
groups$tree

# Plot:
plot(groups)
```

[1]: https://arxiv.org/abs/1905.02928

[2]: https://proceedings.neurips.cc/paper/2020/hash/1819020b02e926785cf3be594d957696-Abstract.html

[3]: https://www.tidymodels.org/find/parsnip/
